---
title: "CSC4008 Assignment1"
author: "Xu Zhihao"
date: "2/7/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Exercise 2.2 (a) - (g) 
(a)
```{r}
age = c(13, 15, 16, 16, 19, 20, 20, 21, 22, 
        22, 25, 25, 25, 25, 30, 33, 33, 35, 
        35, 35, 35, 36, 40, 45, 46, 52, 70)
# mean of the data
mean(age)
# median of the data
median(age)
```
(b)
```{r}
# print the frequency table of the data
table(age)
```
 We can see that this data has 2 modes which is 25 and 35 with frequency 4. Hence this data is bimodal.

(c)
```{r}
# midrange of the data
(13+70)/2
```


(d)
```{r}
summary(age)
```
The first quartile (Q1) is 20.5 and the third quartile (Q3) is 35


(e)
```{r}
# Five number summary
summary(age)[-4]
```

(f)
```{r fig.width=4, fig.height=4}
boxplot(age,main="boxplot of age")
```

(g)
Quantile plot shows each value $x_i$ against with approximately percentage of the data which is lower than or equal to $x_i$. However, the quantile-quantile plot (qq-plot) shows the sample quantile against with a known univariant distribution. Here we show the qq-plot and quantile plot of this data. The known univariant distribution in qq-plot here is normal distribution.
```{r fig.width=8, fig.height=4}
par(mfrow=c(1,2))
freq = as.numeric(table(age))
prop = cumsum(freq)/sum(freq)
plot(prop,unique(age),xlab="f-value",ylab="age",main="Quantile Plot")
qqnorm(age)
qqline(age)
```



## Exercise 2.4 (a)(b)(c)
(a)
```{r}
age = c(23, 23, 27, 27, 39, 41, 47, 49, 50,
        52, 54, 54, 56, 57, 58, 58, 60, 61)
fat = c(9.5, 26.5, 7.8, 17.8, 31.4, 25.9, 27.4, 27.2, 31.2,
        34.6, 42.5, 28.8, 33.4, 30.2, 34.1, 32.9, 41.2, 35.7)
# mean, median, and standard deviation of age
mean(age)
median(age)
sd(age) * sqrt(17/18)
# mean, median, and standard deviation of %fat
mean(fat)
median(fat)
sd(fat) * sqrt(17/18)
```

(b)
```{r fig.width=8, fig.height=4}
par(mfrow=c(1,2))
boxplot(age,main="boxplot of age")
boxplot(fat,main="boxplot of %fat")
```

(c)
```{r fig.width=8, fig.height=4}
par(mfrow=c(1,2))
plot(age,fat,main="scatter plot of age and %fat",ylab="%fat",pch=19)
qqplot(age, fat, main = "Q-Q Plot of Age against %fat",ylab="%fat", pch = 19)
```


## Exercise 2.6 (a)(b)(c)
(a) Euclidean distance
```{r}
a = c(22, 1, 42, 10)
b = c(20, 0, 36, 8)
norm(a-b,type="2")
```

(b) Manhattan distance
```{r}
sum(abs(a-b))
```

(c) Minkowski distance with $q=3$
```{r}
sum((a-b)^3)^{1/3}
```


## Exercise 2.8 (a)(b)
(a)
```{r}
Norm = function(x,type="2"){
  x = as.matrix(x)
  x = x - c(1.4,1.6)
  return(norm(x,type=type))
}
cossim = function(x0){
  x0 = as.numeric(x0)
  x = c(1.4,1.6)
  return(x%*%x0/(norm(x,"2")*norm(x0,"2")))
}
a1 = c(1.5,2,1.6,1.2,1.5)
a2 = c(1.7,1.9,1.8,1.5,1.0)
X = data.frame(a1,a2,row.names=c("x1","x2","x3","x4","x5"))
```

Euclidean distance
```{r}
apply(X,1,Norm,type="2")
# Rank from smallest to largest
rank(apply(X,1,Norm,type="2"))
```

Manhattan distance
```{r}
apply(X,1,Norm,type="1")
# Rank from smallest to largest
rank(apply(X,1,Norm,type="1"))
```

Supremum distance
```{r}
apply(X,1,Norm,type="M")
# Rank from smallest to largest
rank(apply(X,1,Norm,type="M"),ties.method= "min")
```

Cosine similarity
```{r}
apply(X,1,cossim)
# Rank from smallest to largest
rank(apply(X,1,cossim))
```

```{r}
Enorm = function(x,x0){
  x = x - x0
  return(norm(x,"2"))
}
X = t(apply(X, 1, function(x) {x / norm(x,"2")}))
x0 = c(1.4,1.6)
x0 = x0/norm(x0,"2")

apply(X,1,Enorm,x0=x0)
# Rank from smallest to largest
rank(apply(X,1,Enorm,x0=x0))
```


